Hamiltonian ABC

### 2015-03-03 draft.tex
penpenultimate

\vspace{-0.1in}
\subsection{Bayesian Logistic Regression}\label{sec:auto}
We perform Bayesian inference on Bayesian inference on a logistic regression model using the digits $0$ and $1$ from MNIST.  Despite its simplicity, the model still represents a high-dimensional problem for HABC ($D=1568$).  We first run stochastic gradient descent to find $\thetav_{\text{MAP}}$ using the true gradient.  We then run HABC SGLD and SG-Thermostats starting  $\thetav_{\text{MAP}}$ to discover how well the algorithms explore the posterior.  We compare with SGLD and SG-Thermostats using the {\em true} gradients.

We use $n=100$ mini-batches and $R=10$ for SPSA.  Figure~\ref{fig:blr} shows samples projected onto 2 dimensions using random projections.  We can clearly see that the trajectories using SPSA exhibit very similar behavior to Bayesian learning with the true gradients.  
\begin{figure}[t]
%\vskip 0.2in
\begin{center}
\includegraphics[width=0.95\columnwidth]{./images/lr/random_proj.png}
\vspace{-0.15in}
\caption{\small{Bayesian logistic regression sampling trajectories randomly projected.  {\bf White circle:} $\thetav_{\text{MAP}}$}; {\bf Dark blue:} true SGLD; {\bf Red:} SPSA SGLD; {\bf Green:} true SG-Thermostats; {\bf Cyan:} SPSA SG-Thermostats.}
\label{fig:blr}
\end{center}
\vskip -0.2in
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{DISCUSSION AND CONCLUSION} \label{sec:conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Hamiltonian ABC proposes a new set of algorithms for Bayesian inference of likelihood-free models.  HABC builds  upon the connections between Hamilton Monte Carlo with stochastic gradients and well-established gradient approximations based on a minimal number of forward simulations, even in high-dimensions.  We have performed some preliminary experiments showing the feasibility of running ABC on both small and large problems, and we hope that the door has been opened for exploration of larger simulation-based models using HABC.

@@

Expensive simulators are an important class of models that we do not address in this work.  However, previous work in Bayesian inference has shown the usefulness of HMC-based proposals based on Gaussian process of log-likelihood surfaces \cite{rasmussen:2003}.   We could similarly use HABC with recent ABC surrogate models \cite{Meeds2014GpsUai,wilkinson:2014} to avoid simulations whenever possible, yet still  benefit from Hamiltonian dynamics.  
%We address this scenario in Section~\ref{future}.  
%Shown for relatively fast simulators.  Extensions to surrogates with GPS (Meeds2014GpsUai, rasmussen:2003, wilkinson:2014) can use derivative of GP for likelihood as GP of gradient.  These could be matched by occasional gradient estimates (Osborne).

>>>>

\vspace{-0.1in}
\subsection{Bayesian Logistic Regression}\label{sec:auto}
We perform Bayesian inference on Bayesian inference on a logistic regression model using the digits $0$ and $1$ from MNIST.  Despite its simplicity, the model still represents a high-dimensional problem for HABC ($D=1568$).  We first run stochastic gradient descent to find $\thetav_{\text{MAP}}$ using the true gradient.  We then run HABC SGLD and SG-Thermostats starting  $\thetav_{\text{MAP}}$ to discover how well the algorithms explore the posterior.  We compare with SGLD and SG-Thermostats using the {\em true} gradients.   
%
We use $n=100$ mini-batches and $R=10$ for SPSA.  Figure~\ref{fig:blr} shows samples projected onto 2 dimensions using random projections.  We can clearly see that the trajectories using SPSA exhibit very similar behavior to Bayesian learning with the true gradients.  This is very positive result that indicates HABC can successfully exploit the noisy and less informative gradients of SPSA.
\begin{figure}[t]
%\vskip 0.2in
\begin{center}
\includegraphics[width=0.95\columnwidth]{./images/lr/random_proj.png}
\vspace{-0.15in}
\caption{\small{Bayesian logistic regression sampling trajectories randomly projected.  {\bf White circle:} $\thetav_{\text{MAP}}$}; {\bf Dark blue:} true SGLD; {\bf Red:} SPSA SGLD; {\bf Green:} true SG-Thermostats; {\bf Cyan:} SPSA SG-Thermostats.}
\label{fig:blr}
\end{center}
\vskip -0.2in
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{DISCUSSION AND CONCLUSION} \label{sec:conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{-0.1in}
Hamiltonian ABC proposes a new set of algorithms for Bayesian inference of likelihood-free models.  HABC builds  upon the connections between Hamilton Monte Carlo with stochastic gradients and well-established gradient approximations based on a minimal number of forward simulations, even in high-dimensions.  We have performed some preliminary experiments showing the feasibility of running ABC on both small and large problems, and we hope that the door has been opened for exploration of larger simulation-based models using HABC.

@@

Expensive simulators are an important class of models that we do not address in this work.  However, previous work in Bayesian inference has shown the usefulness of HMC-based proposals based on Gaussian process of log-likelihood surfaces \cite{rasmussen:2003}.   We could similarly use HABC with  ABC surrogate models \cite{Meeds2014GpsUai,wilkinson:2014} to minimize simulation calls, yet still  benefit from Hamiltonian dynamics.  
%We address this scenario in Section~\ref{future}.  
%Shown for relatively fast simulators.  Extensions to surrogates with GPS (Meeds2014GpsUai, rasmussen:2003, wilkinson:2014) can use derivative of GP for likelihood as GP of gradient.  These could be matched by occasional gradient estimates (Osborne).
